Precise spiking motifs in neurobiological and neuromorphic data

### 2022-11-08 manuscript.tex
Update on Overleaf.

% follow-ups
Interestingly, the paper by~\citep{izhikevich_polychronization_2006} stirred a lively debate in the field of computational neuroscience, with a general positive acceptance, but relatively few works extended this seminal paper. Indeed, there were already existing models of synaptic delay learning in spiking neural networks, see for instance~\citep{huning_synaptic_1998} or~\citep{eurich_dynamics_1999} yet they had not shown potential applications to the detection of spiking motifs. A popular model for the detection of latency patterns is the \emph{tempotron}~\citep{gutig_tempotron_2006}, particularly reviewed in~\citep{gutig_spike_2014}. The \emph{tempotron} is a supervised synaptic learning algorithm which classifies a distractor from a target motif, in order to extend the perceptron which does not incorporate a spike timing framework. The \emph{tempotron} learning rule is derived by an optimization process and takes the form of a STDP rule. The limits of this model is that its output is only binary and that its storage capacity are limited. The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency~\citep{susi_nmnsd-spiking_2021}. Additional models for detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}. As such, these models do not explicitly take advantage of the representation capacity offered by polychronous groups. 
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{Extensions of Izhikevitch's polychronization model}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}. Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete for the reproduction of all results presented in that manuscript. This more recent work details how this code was slightly corrected. It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process.
%
Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}. Based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PNG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PNG activation. They demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system. Stimulation of the trained network produces the activation of a PNG, i.e. the propagation of firing activity through multiple layers due to convergent patterns of firing.

>>>>

% follow-ups
Interestingly, the paper by~\citep{izhikevich_polychronization_2006} stirred a lively debate in the field of computational neuroscience, with a general positive acceptance, but relatively few works extended this seminal paper. Indeed, there were already existing models of synaptic delay learning in spiking neural networks, see for instance~\citep{huning_synaptic_1998} or~\citep{eurich_dynamics_1999} yet they had not shown potential applications to the detection of spiking motifs. A popular model for the detection of latency patterns is the \emph{tempotron}~\citep{gutig_tempotron_2006}, particularly reviewed in~\citep{gutig_spike_2014}. The \emph{tempotron} is a supervised synaptic learning algorithm which classifies a distractor from a target motif, in order to extend the perceptron which does not incorporate a spike timing framework. The \emph{tempotron} learning rule is derived by an optimization process and takes the form of a STDP rule. The limits of this model is that its output is only binary and that its storage capacity are limited. The recent ``multi-neuronal spike sequence detector'' architecture integrates the weight- and delay-adjustment methods by combining plasticity with the neurocomputational feature spike latency~\citep{susi_nmnsd-spiking_2021}. Additional models for detection of latency patterns are presented in the extensive (graph-centric) review on synchronization in time-varying networks~\citep{ghosh_synchronization_2021,ghosh_synchronized_2022}. As such, these models do not explicitly take advantage of the representation capacity offered by polychronous groups. 
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{Extensions of Izhikevitch's polychronization model}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}. Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete for the reproduction of all results presented in that manuscript. This more recent work details how this code was slightly corrected. It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process.
%
Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}. Based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PNG activation. They demonstrate the use of this method by investigating the claim that PNGs might provide the foundation of a representational system. Stimulation of the trained network produces the activation of a PNG, i.e. the propagation of firing activity through multiple layers due to convergent patterns of firing.
